---
title: "Exe#2_Optical_interconnection_network data set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rsample)
library(caret)
library(tidyverse)
library(dplyr)    
library(ggplot2)  
library(visdat)
library(recipes) 
library(readr)
library(vip)      # for variable importance
```

```{r echo=FALSE}
OIN_data <- read.csv("data/optical_interconnection_network.csv",dec = ",", sep = ";" )
#OIN_data = na.omit(OIN_data)
clean_OIN_data <- OIN_data[,c(1:10)]
head(clean_OIN_data)
set.seed(123)
OIN_data_split <- initial_split(clean_OIN_data, prop = .7)
OIN_data_train <- training(OIN_data_split)
OIN_data_test  <- testing(OIN_data_split)
```

# 1. Depending on the type of response variable, apply a linear or logistic regression model.
## First, apply the model to your data without pre-applying feature engineering processes.
```{r}
model1 <- lm(Channel.Utilization ~ Processor.Utilization  + Channel.Waiting.Time + Input.Waiting.Time + Network.Response.Time,  data = OIN_data_train)
summary(model1)
```

## Visualize multiple regression model

```{r echo=FALSE}
library(moonBook)


require(ggeffects)
require(rgl)
plot(model1)

```

## Create and a apply a blueprint of feature engineering processes that you think will help your model improve.
```{r}
blueprint <- recipe(Channel.Utilization ~ ., data = OIN_data_train) %>%
  step_YeoJohnson(all_numeric()) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE)
```

## Now reapply the model to your data that has been feature engineered.

```{r}
cv <- trainControl(
  method = "repeatedcv", 
  number = 10, 
  repeats = 5
  )
hyper_grid <- expand.grid(k = seq(2, 25, by = 1))

blueprint_model <- train(
  blueprint, 
  data = OIN_data_train, 
  method = "knn", 
  trControl = cv, 
  tuneGrid = hyper_grid,
  metric = "RMSE"
  )
blueprint_model

```

## Did your model performance improve?
##### Yes, The *RMS* become lower than the previous model without applyed *feature engineered*.


# 2. Apply a principal component regression model.
## Perform a grid search over several components.
## Identify and explain the performance of the optimal model.


```{r}
# Build the model on training set
set.seed(123)
cv_pcr <- train(
  Channel.Utilization ~., 
  data = OIN_data_train, 
  method = "pcr",
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
  )
# Plot model RMSE vs different values of components
plot(cv_pcr)
# Print the best tuning parameter ncomp that
# minimize the cross-validation error, RMSE
cv_pcr$bestTune
```

```{r}
# Summarize the final model
summary(cv_pcr$finalModel)
```
```{r}

# Model performance metrics

cv_pcr$results %>%
  filter(ncomp == as.numeric(cv_pcr$bestTune))
```



# 3.Apply a partial least squares regression model.
## Perform a grid search over several components.
##Identify and explain the performance of the optimal model.Apply a partial least squares regression model.
```{r}
set.seed(123)
cv_pls <- train(
  Channel.Utilization ~., 
  data = OIN_data_train, 
  method = "pls",
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
  )
# model with lowest RMSE
cv_pls$bestTune

cv_pls$results %>%
  filter(ncomp == as.numeric(cv_pls$bestTune))

# plot cross-validated RMSE
plot(cv_pls)
```

# 4. Apply a regularized regression model.
## Perform a grid search across alpha parameter values ranging between 0–1.
```{r}
X <- model.matrix(Channel.Utilization ~ ., OIN_data_train)[, -1]
Y <- log(OIN_data_train$Channel.Utilization)

hyper_grid <- expand.grid(
  alpha = seq(0, 1, by = .25),
  lambda = c(0.1, 10, 100, 1000, 10000)
)

set.seed(123)
cv_glmnet <- train(
  x = X,
  y = Y,
  method = "glmnet",
  preProc = c("zv", "center", "scale"),
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)

```

## What is the optimal alpha and lambda values?
## What is the MSE and RMSE for this optimal model?
```{r}
cv_glmnet$results %>%
  filter(
    alpha == cv_glmnet$bestTune$alpha,
    lambda == cv_glmnet$bestTune$lambda
  )
```

## How does it compare to your previous models?
```{r}

pred <- predict(cv_glmnet, X)

# compute RMSE of transformed predicted
RMSE(exp(pred), exp(Y))
```


# 5. Pick the best performing model from above.
## Identify the most influential features for this model.
## Plot the top 10 most influential features.

```{r}
vip(cv_glmnet, num_features = 10, geom = "point") 
```

## Do these features have positive or negative impacts on your response variable?
##### Thefeatures have positive impacts.

